\section{Module 8. Skull stripping}
Module 8 has implemented unit tests with consist of some methods required to skull strip the image. All of below mentioned tests were \textbf{passed}.
Implemented methods:
\begin{itemize}
    \item {\textbf{def test strel(self)} - which tests working of strel method. It was checked input parameters return appropriate output.}
    \item {\textbf{def test binarization(self)} - which tests working of binarization method. It was checked input parameters - image array return appropriate output - binary image.}
    \item {\textbf{def test main8 incorrect input} - which tests working of main8 function. It was checked wrong input parameters return "Unexpected data format in module number 8!"}
    \item {\textbf{def test run preprocessing frist condition} - which tests working of first condition. It was checked input with too small BR or too large ratio return appropriate output.}
    \item {\textbf{def test main8 adding skull stripping mask(self)} - which tests working of main08 function. It was checked if mask are add to mri class.}
    \item {\textbf{def test main08 not changing other inputs(self)} - which tests working of main08 function. It was checked if to mri class is add only skull stripping mask and other fields are the same as input.}
    \item {\textbf{def test main08 mask length(self):} - which tests working of main08 function. It was checked if mask has the same length as input image in mri class.}
\end{itemize}
\begin{lstlisting}[language=Python, caption = Implemented tests examples.]
    @mock.patch("core.inc.module08.SkullStripping.preprocessing")
    def test_run_preprocessing_frist_condition(self, mock_obj):
        mock_obj.return_value = (None, None, None, 10, None)
        ss = SkullStripping(None)
        self.assertEqual(ss.run(), 0)
        mock_obj.return_value = (None, None, None, 100, 2.0)
        self.assertEqual(ss.run(), 0)

    def test_main8_adding_skull_stripping_mask(self):
        self.dwi.skull_stripping_mask = []
        input_skull_stripping_mask = self.dwi.skull_stripping_mask
        output_object = module08.main8(self.dwi)
        self.assertNotEqual(output_object.skull_stripping_mask[:,:,0], input_skull_stripping_mask)

    def test_main08_mask_length(self):
        self.dwi.skull_stripping_mask = numpy.zeros_like(self.dwi.structural_data[:, :, 0])
        input_skull_stripping_mask = self.dwi.skull_stripping_mask
        output_object = module08.main8(self.dwi)
        numpy.testing.assert_equal(len(output_object.skull_stripping_mask), len(input_skull_stripping_mask))
\end{lstlisting}