\section{Tools}

\textbf{Python} - The main programming language of this project is Python. The utilized language version is \textbf{Python 3.5.4} (last "bugfix" release of Python 3.5).

\textbf{VTK} - The Visualization Toolkit (VTK) is an open-source software system for 3D computer graphics, modeling, image processing, volume rendering, scientific visualization, and information visualization.
The project uses version \textbf{7.0}, as it is compatible with python 3.5 (compatibility assured by menpo project\footnote{http://www.menpo.org/}).

\textbf{Cython} - Cython is an optimising static compiler for both the Python programming language and the extended Cython programming language. It allows to represent the most computionally complex operations with more efficient code. This project uses version \textbf{0.26.1}.

\textbf{PyQT} - PyQt brings together the Qt C++ cross-platform application framework and the cross-platform interpreted language Python. This project uses its version \textbf{5.6} to create Graphical User Interface (GUI) for the application.

\textbf{SciPy} - SciPy is a collection of open source packages, which serves to import .mat MATLAB data into NumPy arrays and perform computations on them.

\textbf{Conda} - Conda is an open source package management system and environment management system that runs on Windows, macOS and Linux. Conda quickly installs, runs and updates packages and their dependencies. This project uses \textbf{Anaconda3} distribution to provide standarized enviroment with all of the required packages for our developers.